---
title: "Stat 341 -- HW 1 "
author: "Ben Steves"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document: default
  html_document: default
  word_document: default
---

```{r, setup, include = FALSE, message=FALSE}
# load packages that are going to be used
library(mosaic)      # this loads ggformula (for plotting), etc. too
library(fastR2)      # some data sets
library(pander)      # nicely formatted tables with pander()
library(knitr)       # so you can use kable()
library(patchwork)   # for combining plots

# part of tidyverse for data wrangling
library(dplyr)
library(tidyr)
library(purrr)

# several packages for bayesian stuff -- more to come later
library(rethinking)  # related to text
library(tidybayes)    
library(bayesplot)
library(CalvinBayes)


# Some customization. You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  fig.width = 4,    # adjust this to make figures wider or narrower
  fig.height = 2.5, # adjust this to make figures taller or shorrter
  size = "small")   # slightly smaller font for code
```


<!-- A few math abbreviations -->

\newcommand{\Prob}{\operatorname{Pr}}
\newcommand{\Binom}{\operatorname{Binom}}
\newcommand{\Unif}{\operatorname{Unif}}
\newcommand{\Triangle}{\operatorname{Triangle}}
\newcommand{\Norm}{\operatorname{Norm}}
\newcommand{\Beta}{\operatorname{Beta}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

## 2H1

$P(A) = 0.5$

$P(B) = 0.5$

$P(twins | A) = 0.1$

$P(twins | B) = 0.2$

```{r 2h1}
prior <- c(0.5, 0.5)
likelihood <- c(0.1, 0.2)
posterior <- prior * likelihood
posterior_updated <- posterior/sum(posterior)
prob <- sum(posterior_updated*likelihood)
prob
```

The probability of a random species of panda having two consecutive birthings of twins is `r prob`

## 2H2

```{r 2h2}
prior <- c(0.5, 0.5)
likelihood <- c(0.1, 0.2)
posterior <- prior * likelihood
posterior_updated <- posterior/sum(posterior)
posterior_updated
```

The probability of a panda from species A having two consecutive birthings of twins is `r posterior_updated[1]`

## 2H3

```{r}
#first part - twins
prior1 <- c(0.5, 0.5)
likelihood1 <- c(0.1, 0.2)
posterior1 <- prior1 * likelihood1
posterior_updated1 <- posterior1/sum(posterior1)

#second part - singleton
prior2 <- posterior_updated1
likelihood2 <- c(0.9, 0.8)
posterior2 <- prior2 * likelihood2
posterior_updated2 <- posterior2/sum(posterior2)
posterior_updated2
```

The probability of a panda from species A having a birthing of twins then a singleton is `r posterior_updated2[1]`

## 2H4

```{r}
# with results
prior1 <- c(0.5, 0.5)
likelihood <- c(0.8, 0.65)
posterior1 <- prior1 * likelihood
posterior_updated1 <- posterior1/sum(posterior1)

# add previous data
prior2 <- c(0.36, 0.64) 
posterior2 <- prior2 * likelihood
posterior_updated2 <- posterior2/sum(posterior2)
posterior_updated2
```




The probability of a panda tested as species A is `r posterior_updated2[1]`