---
title: "Stat 341 -- PS 5"
author: "Ben Steves"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document: default
  html_document: default
  word_document: default
---

```{r, setup, include = FALSE, message=FALSE}
# load packages that are going to be used
library(mosaic)      # this loads ggformula (for plotting), etc. too
library(fastR2)      # some data sets
library(pander)      # nicely formatted tables with pander()
library(knitr)       # so you can use kable()
library(patchwork)   # for combining plots

# part of tidyverse for data wrangling
library(dplyr)
library(tidyr)
library(purrr)

# several packages for bayesian stuff -- more to come later
library(rethinking)  # related to text
library(tidybayes)    
library(bayesplot)
library(CalvinBayes)


# Some customization. You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  fig.width = 4,    # adjust this to make figures wider or narrower
  fig.height = 2.5, # adjust this to make figures taller or shorrter
  size = "small")   # slightly smaller font for code
```


<!-- A few math abbreviations -->

\newcommand{\Prob}{\operatorname{Pr}}
\newcommand{\Binom}{\operatorname{Binom}}
\newcommand{\Unif}{\operatorname{Unif}}
\newcommand{\Triangle}{\operatorname{Triangle}}
\newcommand{\Norm}{\operatorname{Norm}}
\newcommand{\Beta}{\operatorname{Beta}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

## 4E1

Of the three, the first one is the likelihood - $y_i \sim Normal(\mu, \sigma)$

## 4E2

There are two parameters in the posterior, $\mu$ and $\sigma$

## 4M1

```{r}
Prior4m1 <- 
  tibble(
    mu =     rnorm(1e4, 0, 10),
    sigma =  rexp(1e4, 1),
    y = rnorm(1e4, mu, sigma)
  )
head(Prior4m1)

```

## 4M2

```{r}
Prior4m2 <- 
    alist(
      height ~ dnorm(mu, sigma),
      mu     ~ dnorm(0, 10),
      sigma  ~ dexp(1)
    )
Prior4m2
```

## 4M3

$y_i \sim Normal(\mu, \sigma)$

$\mu_i \sim \alpha + \beta x_i$

$\alpha \sim Normal(0, 10)$

$\beta \sim Uniform(0, 1)$

$\sigma \sim Exponential(1)$

## 6.1

```{r}
set.seed(145)
mean <- log(2)

f <- function(sigma) {
  (exp((2*mean)+sigma^2)*(exp(sigma^2)-1)) - 25
}

# put in endpoints to help generate zero
sd <- uniroot(f, c(0.4, 6))$root
prior6.1 <- rlnorm(n=1e5, mean, sd)
median(prior6.1)

# throw in a data frame to show values
sd(prior6.1)
logparams <- c(mean = mean, sd = sd); logparams
```

$Y \sim Lognormal(0.693, 1.056)$ has a median of 2 and a standard deviation of 5. 
