---
title: "Stat 341 -- PS 6"
author: "Ben Steves"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document: default
  html_document: default
  word_document: default
---

```{r, setup, include = FALSE, message=FALSE}
# load packages that are going to be used
library(mosaic)      # this loads ggformula (for plotting), etc. too
library(fastR2)      # some data sets
library(pander)      # nicely formatted tables with pander()
library(knitr)       # so you can use kable()
library(patchwork)   # for combining plots

# part of tidyverse for data wrangling
library(dplyr)
library(tidyr)
library(purrr)

# several packages for bayesian stuff -- more to come later
library(rethinking)  # related to text
library(tidybayes)    
library(bayesplot)
library(CalvinBayes)


# Some customization. You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  fig.width = 4,    # adjust this to make figures wider or narrower
  fig.height = 2.5, # adjust this to make figures taller or shorrter
  size = "small")   # slightly smaller font for code
```


<!-- A few math abbreviations -->

\newcommand{\Prob}{\operatorname{Pr}}
\newcommand{\Binom}{\operatorname{Binom}}
\newcommand{\Unif}{\operatorname{Unif}}
\newcommand{\Triangle}{\operatorname{Triangle}}
\newcommand{\Norm}{\operatorname{Norm}}
\newcommand{\Beta}{\operatorname{Beta}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

## 4H1

```{r}
# filter adults
data(Howell1)
HowellAdults <-
  Howell1 %>%
  filter(age >= 18)
```

```{r}
# fit model
model4H1 <- 
  quap(
    data = HowellAdults,
    alist(
      height ~ dnorm(mu, sigma),
      mu    <- alpha_0 + beta_1 * (weight - mean(weight)),
      alpha_0 ~ dnorm(178, 20),
      beta_1 ~ dlnorm(0, 1),
      sigma  ~ dunif(0, 50)
    )
  )
# get posteriors
mean_weight = mean( ~ weight, data = HowellAdults)
Posterior4H1 <-
  extract.samples(model4H1, n = 1000) %>%
  mutate(beta_0 = alpha_0 - beta_1 * mean_weight)
```

```{r}
# using function from notes
ind_height <- function(post, weight, prob = 0.89) {
  n <- nrow(post)
  post %>%
    mutate(ind_height = rnorm(n, beta_0 + beta_1 * weight, sigma)) %>%
    mean_hdi(ind_height, .width = prob) %>%
    mutate(weight = weight)
}

# calculate HDI
weights <- c(46.95, 43.72, 64.78, 32.59, 54.63)
map_dfr(weights, ~ ind_height(Posterior4H1, .x))
```


## 4H2

### a)

```{r}
# filter kids
HowellKids <-
  Howell1 %>%
  filter(age < 18)
```

```{r}
# fit model. prior reflects possible kids height.
model4H2 <- 
  quap(
    data = HowellKids,
    alist(
      height ~ dnorm(mu, sigma),
      mu    <- alpha_0 + beta_1 * (weight - mean(weight)),
      alpha_0 ~ dnorm(115, 30),
      beta_1 ~ dlnorm(0, 1),
      sigma  ~ dunif(0, 50)
    )
  )

# change in height per 10 units of weight
summary4H2 <- precis(model4H2); summary4H2
summary4H2$mean[2]*10
```

There is a 2.72 cm increase in height for each additional unit of weight. To get $\beta_1$ for each 10 units of weight, it's just 27.2, or 10*2.72. The height for a child of average weight is 108.32. The standard deviation of heights is 8.44.

### b)

```{r fig.width = 7, fig.height = 4}
# mean and individual avg
Link4H2 <-
  link(model4H2, tibble(weight = seq(2, 45, by = 1)))
Sim4H2 <- 
  sim(model4H2, tibble(weight = seq(2, 45, by = 1)))

Avg4H2 <- 
  apply(Link4H2, 2, HPDI) 

Avg4H2df <- as.data.frame(t(Avg4H2)) %>%
    mutate(weight = seq(2, 45, by = 1),
           ymin = `|0.89`,
           ymax = `0.89|`)

Ind4H2 <- 
  apply(Sim4H2, 2, HPDI)

Ind4H2df <- as.data.frame(t(Ind4H2)) %>%
    mutate(weight = seq(2, 45, by = 1),
           ymin = `|0.89`,
           ymax = `0.89|`)

gf_ribbon(ymin + ymax ~ weight, data = Ind4H2df, fill = ~"ind") %>%
  gf_ribbon(ymin + ymax ~ weight, data = Avg4H2df, fill = ~"avg") %>%
  gf_point(height ~ weight, data = HowellKids, inherit = FALSE) 
```

### c)

The model is concerning for a few reasons. For one, the model is underestimates the actual heights of kids for given weight. In that same regard, there are many points that lie over the HDI boundary, mostly in the weight range of 15-35. The estimations were made assuming that the relationship between height and weight in kids is linear. However, this relationship is not very linear and is more quadratic. $\beta_1$ being log-normal is bad as it assumes a linear shape for the slope, so the distribution type for $\beta_1$ should be something different, something to fit a polynomial regression. 
