---
title: "Stat 341 -- PS 13"
author: "Ben Steves"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document: default
  html_document: default
  word_document: default
---

```{r, setup, include = FALSE, message=FALSE}
# load packages that are going to be used
library(mosaic)      # this loads ggformula (for plotting), etc. too
library(fastR2)      # some data sets
library(pander)      # nicely formatted tables with pander()
library(knitr)       # so you can use kable()
library(patchwork)   # for combining plots

# part of tidyverse for data wrangling
library(dplyr)
library(tidyr)
library(purrr)

# several packages for bayesian stuff -- more to come later
library(rethinking)  # related to text
library(tidybayes)    
library(bayesplot)
library(CalvinBayes)


# Some customization. You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  fig.width = 4,    # adjust this to make figures wider or narrower
  fig.height = 2.5, # adjust this to make figures taller or shorrter
  size = "small")   # slightly smaller font for code
```


<!-- A few math abbreviations -->

\newcommand{\Prob}{\operatorname{Pr}}
\newcommand{\Binom}{\operatorname{Binom}}
\newcommand{\Unif}{\operatorname{Unif}}
\newcommand{\Triangle}{\operatorname{Triangle}}
\newcommand{\Norm}{\operatorname{Norm}}
\newcommand{\Beta}{\operatorname{Beta}}
\newcommand{\E}{\operatorname{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\SD}{\operatorname{SD}}

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

## 8E1

1. Bread rise ~ Yeast:Temperature. Colder baking temperatures have an effect on the relationship between yeast and the bread rising

2: Income ~ Education level:Gender. Females usually earn less than males.

3: Car works ~ Gasoline:Year built. The age of a car may have an impact on the functionality of a car. 

## 8M2

```{r}
# just to look at variables
data(tulips, package = "rethinking")
```

$\mu_i = \beta_0 + \beta_w w_i + \beta_s s_i + \beta_t t_i + \beta_{sw} s_iw_i + \beta_{st} s_it_i + \beta_{wt} w_it_i + \beta_{wst} w_is_it_i$

To get a bloom size of 0, make all values for $t_i$ in the equation equal to 1. 

$\mu_i = \beta_0 + \beta_w w_i + \beta_s s_i + \beta_t + \beta_{sw} s_iw_i + \beta_{st} s_i + \beta_{wt} w_i + \beta_{wst} w_is_i$
